---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.



library(classdata)
data("ames")
head(ames)

Step 1 Result (Abhi's Work):

As a team, we found the variables for the dataset:
          Variable          Type          Meaning                                             Expected Range
1.        Parcel ID         Character     Unique property identifier                          Unique values
2.        Address           Character     Property address                                    Text
3.        Style             Character     Building style                                      Categories
4.        Occupancy         Character     Type of occupancy                                   Categories
5.        Sale Date         Date          Date property was sold                              2017–2024+
6.        Sale Price        Numeric       Final sale price in dollars                         0 – 1,000,000+
7.        Multi Sale        Character     Whether multiple sales occurred                     NA/Yes
8.        YearBuilt         Numeric       Year property was built                             1900–2024
9.        Acres             Numeric       Lot size in Acres                                   0–5+
10.       TotalLivingArea   Numeric       Total above-ground living area in square feet       500–5000+
11.       Bedrooms          Numeric       Number of bedrooms                                  0-6
12.       FinishedBsmtArea  Numeric       Finished basement area in square feet               0–3000
13.       LotArea           Numeric       Lot size in square feet                             1,000–50,000+
14.       AC                Character     Whether home has air conditioning                   Yes/No
15.       FirePlace         Character     Whether home has fireplace                          Yes/No
16.       Neighborhood      Character     Neighborhood classification                         Categories


- Step 2 (Yash's work): We choose "Sale Price" as the main variable for our work.

- Step 3 (Kate's work): The range is max - min which is 20500000 - 0 = 20500000. The data is skewed right with 2 significant outliers at ~2mil and ~1.5mil. However, most of the data falls below 200k.
```{r}
library(classdata)
data("ames")
library(ggplot2)
library(scales)

ggplot(ames, aes(x = `Sale Price`)) +
  geom_histogram(binwidth = 50000, color = "black", fill = "gray") +
  coord_cartesian(xlim = c(0, 600000)) +
  scale_x_continuous(labels = label_dollar()) +
  labs(
    title = "Distribution of Ames Home Sale Prices",
    x = "Sale Price",
    y = "Number of Homes"
  )
```
```{r}
sum(ames$`Sale Price` == 0)
```
- Step 3 continuation (Yash's work): Upon plotting the histogram, I noticed there were 2206 observations at 0 dollars for sale price. These likely indicate placeholder, or missing values since houses are never sold for free. Extreme outliers such as prices around 15-21 million dollars are not shown in the plot to better visualize the overall distribution of the sale prices, and to make the plot readable. After doing so, the distribution appears right skewed, with most homes selling between approximately 100000 and 300000 dollars respectively, with the most amount of houses selling at 200000 dollars. 

Step 4:  
- Abhi's work:  
  
I chose **TotalLivingArea (sf)** as a variable that may be related to the main variable, **Sale Price**, because larger homes are generally expected to sell for higher prices.  
  
**Range of TotalLivingArea (sf)**. 
```{r}
living_area <- ames$`TotalLivingArea (sf)`

min(living_area, na.rm = TRUE)
max(living_area, na.rm = TRUE)
```
From the output, the minimum is 0 square feet, the maximum is 6007 square feet, so the range would be 0 square feet to 6007 square feet. The minimum value of 0 sq ft is unusual and likely represents missing or incorrectly entered data.  
  
**Distribution of TotalLivingArea (sf)**. 
```{r}
library(ggplot2)

ggplot(ames, aes(x = `TotalLivingArea (sf)`)) +
  geom_histogram(binwidth = 250, color = "black", fill = "lightblue") +
  labs(
    title = "Distribution of Total Living Area",
    x = "Total Living Area (sq ft)",
    y = "Count"
  )
```
  
The distribution is right-skewed. Most homes fall between 1,000 and 2,500 square feet. 
A small number of homes are very large (above 4,000 sq ft). 
There are also a few values near 0 sq ft, which appear suspicious.
  
**Relationship Between TotalLivingArea and Sale Price**. 
```{r}
library(classdata)
library(dplyr)
library(ggplot2)

data("ames")

ames_clean <- ames %>%
  filter(`Sale Price` > 0,
         `Sale Price` < 1000000,
         `TotalLivingArea (sf)` > 0)

ggplot(ames_clean, aes(x = `TotalLivingArea (sf)`, y = `Sale Price`)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Sale Price vs Total Living Area (Cleaned Data)",
    x = "Total Living Area (sq ft)",
    y = "Sale Price ($)"
  )
```
  
After removing homes with Sale Price equal to 0 and extreme outliers above $1,000,000, the relationship between Total Living Area and Sale Price shows a clear positive linear trend. Larger homes generally sell for higher prices. Most homes fall between 1,000 and 2,500 square feet and sell between $100,000 and $400,000. There is greater variability in sale price among larger homes. The very large homes help explain the extreme high sale price outliers observed in Step 3, as properties with greater square footage tend to command much higher prices and contribute to the right-skew. However, this variable does not explain the $0 sale prices, which likely represent missing or incorrectly recorded data.
