---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

Step 1 Result (Abhi's Work):

```{r}
library(classdata)
data("ames")
head(ames)
```

```{r}
library(tibble)
library(knitr)

step1_table <- tibble(
  Variable = c(
    "Parcel ID",
    "Address",
    "Style",
    "Occupancy",
    "Sale Date",
    "Sale Price",
    "Multi Sale",
    "YearBuilt",
    "Acres",
    "TotalLivingArea (sf)",
    "Bedrooms",
    "FinishedBsmtArea (sf)",
    "LotArea (sf)",
    "AC",
    "FirePlace",
    "Neighborhood"
  ),
  Type = c(
    "Character",
    "Character",
    "Character",
    "Character",
    "Date",
    "Numeric",
    "Character",
    "Numeric",
    "Numeric",
    "Numeric",
    "Numeric",
    "Numeric",
    "Numeric",
    "Character",
    "Character",
    "Character"
  ),
  Meaning = c(
    "Unique property identifier",
    "Property address",
    "Building style",
    "Type of occupancy",
    "Date property was sold",
    "Final sale price in dollars",
    "Whether multiple sales occurred",
    "Year property was built",
    "Lot size in acres",
    "Total above-ground living area in square feet",
    "Number of bedrooms",
    "Finished basement area in square feet",
    "Lot size in square feet",
    "Whether home has air conditioning",
    "Whether home has fireplace",
    "Neighborhood classification"
  ),
  Expected_Range = c(
    "Unique values",
    "Text",
    "Categories",
    "Categories",
    "2017–2024+",
    "0–1,000,000+",
    "NA/Yes",
    "1900–2024",
    "0–5+",
    "500–5000+",
    "0–6",
    "0–3000",
    "1,000–50,000+",
    "Yes/No",
    "Yes/No",
    "Categories"
  )
)

kable(step1_table, caption = "Step 1: Dataset Variables Overview")
```

- Step 2 (Yash's work): We choose "Sale Price" as the main variable for our work.

- Step 3 (Kate's work): The range is max - min which is 20500000 - 0 = 20500000. The data is skewed right with 2 significant outliers at ~2mil and ~1.5mil. However, most of the data falls below 200k.
```{r}
library(classdata)
data("ames")

library(ggplot2)
library(scales)

ggplot(ames, aes(x = `Sale Price`)) +
  geom_histogram(binwidth = 50000, color = "black", fill = "gray") +
  coord_cartesian(xlim = c(0, 600000)) +
  scale_x_continuous(labels = label_dollar()) +
  labs(
    title = "Distribution of Ames Home Sale Prices",
    x = "Sale Price",
    y = "Number of Homes"
  )
```
```{r}
sum(ames$`Sale Price` == 0)
```
- Step 3 continuation (Yash's work): Upon plotting the histogram, I noticed there were 2206 observations at 0 dollars for sale price. These likely indicate placeholder, or missing values since houses are never sold for free. Extreme outliers such as prices around 15-21 million dollars are not shown in the plot to better visualize the overall distribution of the sale prices, and to make the plot readable. After doing so, the distribution appears right skewed, with most homes selling between approximately 100000 and 300000 dollars respectively, with the most amount of houses selling at 200000 dollars. 

Step 4:  
- Abhi's work:  
  
I chose **TotalLivingArea (sf)** as a variable that may be related to the main variable, **Sale Price**, because larger homes are generally expected to sell for higher prices.  
  
**Range of TotalLivingArea (sf)**. 
```{r}
living_area <- ames$`TotalLivingArea (sf)`

min(living_area, na.rm = TRUE)
max(living_area, na.rm = TRUE)
```
From the output, the minimum is 0 square feet, the maximum is 6007 square feet, so the range would be 0 square feet to 6007 square feet. The minimum value of 0 sq ft is unusual and likely represents missing or incorrectly entered data.  
  
**Distribution of TotalLivingArea (sf)**. 
```{r}

library(ggplot2)

ggplot(ames, aes(x = `TotalLivingArea (sf)`)) +
  geom_histogram(binwidth = 250, color = "black", fill = "lightblue") +
  labs(
    title = "Distribution of Total Living Area",
    x = "Total Living Area (sq ft)",
    y = "Count"
  )
```

The distribution is right-skewed. Most homes fall between 1,000 and 2,500 square feet. 
A small number of homes are very large (above 4,000 sq ft). 
There are also a few values near 0 sq ft, which appear suspicious.
  
**Relationship Between TotalLivingArea and Sale Price**. 
```{r}
library(classdata)
library(dplyr)
library(ggplot2)

data("ames")

ames_clean <- ames %>%
  filter(`Sale Price` > 0,
         `Sale Price` < 1000000,
         `TotalLivingArea (sf)` > 0)

ggplot(ames_clean, aes(x = `TotalLivingArea (sf)`, y = `Sale Price`)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    title = "Sale Price vs Total Living Area (Cleaned Data)",
    x = "Total Living Area (sq ft)",
    y = "Sale Price ($)"
  )
```
  
After removing homes with Sale Price equal to 0 and extreme outliers above $1,000,000, the relationship between Total Living Area and Sale Price shows a clear positive linear trend. Larger homes generally sell for higher prices. Most homes fall between 1,000 and 2,500 square feet and sell between $100,000 and $400,000. There is greater variability in sale price among larger homes. The very large homes help explain the extreme high sale price outliers observed in Step 3, as properties with greater square footage tend to command much higher prices and contribute to the right-skew. However, this variable does not explain the $0 sale prices, which likely represent missing or incorrectly recorded data.






Keaton's Work: 

 The range of the variables Acres spans from 0.000 to 12.012, making the total range 12.012. It is interesting to have a home with 0 acres so there must be missing information.
```{r, message=FALSE, warning=FALSE}
hist(ames$Acres[ames$Acres <= 2],
     breaks = seq(0, 2, by = 0.05),
     main = "Distribution of Lot Size (Acres)",
     xlab = "Acres",
     col = "lightblue")

library(dplyr)
library(ggplot2)

ames_small <- ames %>%
  filter(`Sale Price` > 0,
         `Sale Price` < 600000,
         Acres < 2)

ggplot(ames_small, aes(x = Acres, y = `Sale Price`)) +
  geom_point(alpha = 0.35, size = 1.2) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Sale Price vs Lot Size (Acres)",
       x = "Acres",
       y = "Sale Price ($)") +
  theme_minimal()

```

Explanation: For the variable Acres, it is heavily right skewed. Almost every piece of data in the data set comes from under 2 acres. Within that almost every point is within 0.5 acres, with a long but sparse tail of high amounts of acres. When looking at the relationship between acres and sale price the correlation is weak. This is uprising because it would be expected for the price to increase as the amount of acres associated with the home. The relationship is weak due to the high volume of lower priced homes on limited land. When looking at the oddities found in step 3, the variable of acres can fix to highlight some. Homes that are on large pieces of can be the reasoning for why the price is so large. That the sale price is most related to the amount of land versus the actual home. While the acreage can account for some of the high outlines it does not account for the high consistent of lower values. 

  
