---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.



library(classdata)
data("ames")
head(ames)

Step 1 Result (Abhi's Work):

As a team, we found the variables for the dataset:
          Variable          Type          Meaning                                             Expected Range
1.        Parcel ID         Character     Unique property identifier                          Unique values
2.        Address           Character     Property address                                    Text
3.        Style             Character     Building style                                      Categories
4.        Occupancy         Character     Type of occupancy                                   Categories
5.        Sale Date         Date          Date property was sold                              2017–2024+
6.        Sale Price        Numeric       Final sale price in dollars                         0 – 1,000,000+
7.        Multi Sale        Character     Whether multiple sales occurred                     NA/Yes
8.        YearBuilt         Numeric       Year property was built                             1900–2024
9.        Acres             Numeric       Lot size in Acres                                   0–5+
10.       TotalLivingArea   Numeric       Total above-ground living area in square feet       500–5000+
11.       Bedrooms          Numeric       Number of bedrooms                                  0-6
12.       FinishedBsmtArea  Numeric       Finished basement area in square feet               0–3000
13.       LotArea           Numeric       Lot size in square feet                             1,000–50,000+
14.       AC                Character     Whether home has air conditioning                   Yes/No
15.       FirePlace         Character     Whether home has fireplace                          Yes/No
16.       Neighborhood      Character     Neighborhood classification                         Categories


- Step 2 (Yash's work): We choose "Sale Price" as the main variable for our work.

- Step 3 (Kate's work): The range is max - min which is 20500000 - 0 = 20500000. The data is skewed right with 2 significant outliers at ~2mil and ~1.5mil. However, most of the data falls below 200k.
```{r}
library(classdata)
data("ames")
library(ggplot2)

salePrice <- ames$`Sale Price`
range <- max(salePrice) - min(salePrice)

ggplot(ames, aes(x = salePrice)) +
  geom_histogram(binwidth = 500000, color = "black", fill = "gray")

```


Keaton's Work: 

 The range of the variables Acres spans from 0.000 to 12.012, making the total range 12.012.
```{r, message=FALSE, warning=FALSE}
hist(ames$Acres,
     breaks = seq(0, 15, by = 0.1),
     main = "Histogram of Lot Size (Acres)",
     xlab = "Acres",
     col = "lightblue",
     xlim = c(0, 5))

library(ggplot2)

ggplot(ames, aes(x = Acres, y = salePrice)) +
  geom_point(alpha = 0.25, size = 1) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  coord_cartesian(xlim = c(0, 2), ylim = c(0, 600000)) +
  labs(title = "Sale Price vs Lot Size (Acres)",
       x = "Acres",
       y = "Sale Price ($)") +
  theme_minimal()



```

Explanation: For the variable Acres, it is heavily right skewed. Almost every piece of data in the data set comes from under 3 acres. Within that almost every point is within 0.5 acres, with a long but sparse tail of high amounts of acres. When looking at the relationship between acres and sale price the correlation is weak. This is uprising because it would be expected for the price to increase as the amount of acres associated with the home. The relationship is weak due to the high volume of lower priced homes on limited land. When looking at the oddities found in step 3, the variable of acres can fix to highlight some. Homes that are on large pieces of can be the reasoning for why the price is so large. That the sale price is most related to the amount of land versus the actual home. While the acreage can account for some of the high outlines it does not account for the high consistent of lower values. 
